r5r: rapid realistic routing on multimodal transport networks with R5 in R

### 2020-10-22 r-package/paper2020/paper2020.Rmd
paper updates and edits

r5r has low data requirements, enabling routing analysis on different contexts 
wherever there is open data on street networks from OpenStreetMap (OSM). The 
package is also easily scalable allowing for fast computations at either the 
city or country level analysis. For multimodal transport routing, r5r will 
automatically combine OSM data with one or multiple public transport data sets 
provided by the user in  the standard GTFS format to create a routable transport
network.

@@

- `travel_time_matrix()`: used to compute travel time estimates between one or 
multiple origin/destination pairs for a single departure time or for multiple
departure times over a time window. The output comprises the travel time between
each origin/destination pair. If the user passes a `time_window` parameter
to the function, r5r will calculate multiple travel time matrices departing each 
minute and return additional information on the percentile distribution of travel times
that allows one to grasp service reliability issues.

- `detailed_itineraries()`: used to compute detailed information on either the
shortest or multiple alternative routes between one or multiple origin/destination
pairs for a single departure time. The output includes detailed information on
route alternatives such as the transport mode, waiting time, travel time and 
distance of each segment of the trip.

Both routing functions are versatile so users can easily set customized inputs
such as transport mode, departure date and times, walking and cycling speeds, 
maximum trip duration, walking distances and number of public transport 
transfers, etc.

@@

The first step when using r5r is to set the amount of memory available to Java. 
By default `R` allocates only 512 MB of memory for Java processes, which is not 
enough for large queries using r5r. To increase available memory to 2GB, for 
example, we need to set the `java.parameters` option at the beginning of the 
script as follows:

@@

```{r}
data_path <- system.file("extdata/poa", package = "r5r")
list.files(data_path)

points <- fread(system.file("extdata/poa/poa_hexgrid.csv", package = "r5r"))
points <- points[c(sample(1:nrow(points), 100))]
head(points)
```

@@

To build a routable transport network with r5r, the user only needs one command
line pointing to `setup_r5()` where the input data is stored:

@@

The presence of a `.pbf` file with OpenStreetMap data is mandatory because it is
required to create a street network. Meanwhile, a `GTFS.zip` file is only used to
calculate public transport trips, thus may not be necessary if that's not intended.
Multiple GTFS can be used simultaneously, in which case all files are merged into
a single one automatically by R<sup>5</sup>.

@@

Had `shortest_path` been set to `TRUE`, only the shortest path would have been 
returned:

```{r, message = FALSE}
dit_short <- detailed_itineraries(r5r_core,
                                  origins,
                                  destinations,
                                  mode,
                                  departure_datetime,
                                  max_walk_dist,
                                  shortest_path = TRUE)

setDT(copy(dit_short))[, .(n_itineraries = uniqueN(option)), by = .(fromId, toId)]
```

@@

The [R<sup>5</sup> routing engine](https://github.com/conveyal/r5) is developed 
at [Conveyal](https://www.conveyal.com/) with contributions from several people. 
This work was supported by the Brazilian Institute for Applied Economic Research 
(Ipea).

>>>>

r5r has low data requirements, enabling routing analysis on different contexts 
wherever data on street networks from OpenStreetMap (OSM) is available. The 
package is also easily scalable allowing for fast computations at either the 
city or country level analysis. For multimodal transport routing, r5r will 
automatically combine OSM data with one or multiple public transport data sets 
provided by the user in  the standard GTFS format to create a routable transport
network.

@@

- `travel_time_matrix()`: computes travel time estimates between one or 
multiple origin/destination pairs for a single departure time or for multiple
departure times over a time window. The travel_time_matrix function uses an R5-specific extension to the RAPTOR routing algorithm developed by Conway et al. [@conway2017evidencebased], which provides an efficient and systematic sample of 
multiple departures per minute over a `time_window` set by the user. If the user
passes a `time_window` parameter to the function, r5r will calculate multiple travel 
time matrices departing each minute and return additional information on the percentile distribution of travel times that allows one to grasp service reliability issues.

- `detailed_itineraries()`: computes detailed information on either the
shortest or multiple alternative routes between one or multiple origin/destination
pairs for a single departure time. The output includes detailed information on
route alternatives such as the transport mode, waiting time, travel time and 
distance of each segment of the trip. This function uses an R5-specific extension
to the McRAPTOR routing algorithm to find paths that are optimal or less than
optimal. The specific extension to McRAPTOR to do suboptimal path routing are not documented yet, but a detailed description of base McRAPTOR can be found in Delling et al (2015).

#> incluir essa ref
#> - Delling, D., Pajor, T., & Werneck, R. F. (2015). Round-based public transit routing. Transportation Science, 49(3), 591-604.

Both routing functions are versatile so users can easily set customized inputs
such as transport modes, departure date and times, walking and cycling speeds, 
maximum trip duration, walking distances and number of public transport 
transfers, etc.

@@

The first step when using r5r is to set the amount of memory available to Java. 
By default `R` allocates only 512 MB of memory for Java processes~~, which is not 
enough for large queries using r5r~~. To increase available memory to 2GB, for 
example, we need to set the `java.parameters` option at the beginning of the 
script as follows:

@@

```{r, include = FALSE}

# point to directory with data
data_path <- system.file("extdata/spo", package = "r5r")

```

```{r}
# list files in directory
list.files(data_path)

# read points of origin and destination
points <- fread(file.path(data_path, "spo_hexgrid.csv"))
points <- points[c(sample(1:nrow(points), 100))]
head(points)
```

@@

To build a routable transport network with r5r, the user only needs one command
line pointing to where the input data is stored:

@@

The presence of a `.pbf` file with OpenStreetMap data is mandatory because it is
required to create a street network. Meanwhile, a `GTFS.zip` file is only used to
calculate public transport trips, thus may not be necessary if that's not intended.
Multiple GTFS feeds can be used simultaneously, in which case R<sup>5</sup> will
merge all files into a single network automatically.

@@

@@

The [R<sup>5</sup> routing engine](https://github.com/conveyal/r5) is developed 
at [Conveyal](https://www.conveyal.com/) with contributions from several developers. 
This work was supported by the Brazilian Institute for Applied Economic Research 
(Ipea).
